Characterizing Signal Loss, Error, and Bias in the 21cm Reionization \\Power Spectrum: A Revised Study of PAPER-64

### 2018-06-21 PAPER.tex
small edits until 3.3

The previously best published 21\,cm upper limit result from \citetalias{ali_et_al2015} uses $124$ nights of data to place a $2\sigma$ upper limit 
on $\Delta^{2}(k)$, defined as

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_data.pdf}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data.pdf}
	\caption{A power spectrum of a subset of PAPER-64 data illustrating the use of empirical inverse covariance weighting. The solid red curve is the $2\sigma$ upper limit on the EoR signal estimated from our signal injection framework using empirical inverse covariance weighting. Shown for comparison is the lossy limit prior to signal loss estimation (dashed red). The theoretical $2\sigma$ thermal noise level prediction based on observational parameters is in green, whose calculation is detailed in Section \ref{sec:PSSense}. Additionally, the power spectrum results for the uniform weighted case is shown in three different ways: power spectrum values (black and gray points as positive and negative values, respectively, with $2\sigma$ error bars from bootstrapping), the $2\sigma$ upper limit on the EoR signal using our full signal injection framework (solid blue), and the measured power spectrum values with $2\sigma$ thermal noise errors (gray shaded regions). The vertical dashed black lines signify the horizon limit for this analysis using $30$\,m baselines. In this example, we see that the lossy power spectrum limit is $\sim 4$ orders of magnitude too low when using empirical inverse covariance weighting.}
\label{fig:ps2_data}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_data_add.pdf}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data_add.pdf}
	\caption{A power spectrum of a subset of PAPER-64 data illustrating the use of $\widehat{\textbf{C}}_{\rm eff}$ to minimize signal loss. The solid red curve is the $2\sigma$ upper limit on the EoR signal estimated from our signal injection framework. The theoretical $2\sigma$ thermal noise level prediction based on observational parameters is in green. Additionally, the power spectrum results for the uniform weighted case is shown in three different ways: power spectrum values (black and gray points as positive and negative values, respectively, with $2\sigma$ error bars from bootstrapping), the $2\sigma$ upper limit on the EoR signal using our full signal injection framework (solid blue), and the measured power spectrum values with $2\sigma$ thermal noise errors (gray shaded regions). The vertical dashed black lines signify the horizon limit for this analysis using $30$\,m baselines. This power spectrum result does not use the full dataset's sensitivity as in \citetalias{ali_et_al2015} and Kolopanis et al. (\textit{in prep.}), though we include all analysis changes which have mostly stemmed from revisions regarding signal 
loss, bootstrapping, and the theoretical error computation. We see that the regularization scheme used here produces limits similar to the unweighted limits.}
	\label{fig:ps1_data}
\end{figure*}

@@

In this section we discuss the ways in which we estimate errors for PAPER-64 power spectra. We first walk through an expression for a theoretical error estimation (of thermal noise) based on observational parameters. Although a theoretical model often 
differs from true errors as explained in Section \ref{sec:ErrorOverview}, it is helpful to understand the ideal case and the factors 
that affect its sensitivity. Additionally, we build on the lessons learned about bootstrapping in Section \ref{sec:ErrorOverview} to 
revise our bootstrapping method as applied to PAPER-64 data in order to compute accurate errors from the data itself.

@@

This is lower than the $T_{\rm sys}$ of $500$\,K used in \citetalias{ali_et_al2015} because of several small mis-calculation errors that were 

@@

N_{\rm gps}}{2}}$\Big). A revised \citetalias{ali_et_al2015} analysis should therefore use $N_{\rm bls} \sim 32$ instead of $51$, and this change is taken into account in Figure \ref{fig:sense_check}. However, the analysis in this paper and in Kolopanis et al. (\textit{in prep.}) does not average baselines into groups ($N_{\rm gps} = 1$). For the subset of data presented in this paper, $N_{\rm bls} = 10$.

@@

An additional factor of $\sqrt{2}$ is gained in sensitivity when folding together positive and negative $k$'s.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/sense_check.pdf}
	\caption{An updated prediction for the thermal noise level of PAPER-64 data (black) is shown in comparison to previously 
published sensitivity limits (gray), both computed for the analysis methods used in \citetalias{ali_et_al2015}. Major factors that contribute to the discrepancy are $
\Omega_{\rm eff}$, $N_{\rm days}$ and $N_{\rm bls}$, as in Equation \eqref{eq:sense} and described in Section \ref{sec:PSSense}, which when combined decreases our 
sensitivity (higher noise floor) by a factor of $\sim7$ in mK$^{2}$.}
	\label{fig:sense_check}
\end{figure}

@@

\noindent where $\Delta\nu$ is channel spacing, $\Delta t$ is integration time, $N_{\rm days}$ is the number of daily counts for a 
particular time and frequency that went into our LST-binned set, and $N_{\rm pols}$ is the number of polarizations ($2$ for Stokes 
I). This temperature sets the variance of the Gaussian random noise.

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_noise_add.pdf}
	\includegraphics[width=0.4\textwidth]{plots/ps2_noise_add.pdf}
	\caption{The power spectrum for a noise simulation that mimics the noise level of a subset of PAPER-64 data, where the solid red curve is the $2\sigma$ upper limit on the EoR signal estimated from our signal injection framework using $\widehat{\textbf{C}}_{\rm eff}$. The theoretical $2\sigma$ thermal noise level prediction based on observational parameters (calculated by Equation \eqref{eq:sense}) is in green. Additionally, the power spectrum results for the uniform weighted case is shown in three different ways: power spectrum values (black and gray points as positive and negative values, respectively, with $2\sigma$ error bars), the $2\sigma$ upper limit on the EoR signal using our full signal injection framework (solid blue), and the measured power spectrum values with $2\sigma$ thermal noise errors (gray shaded regions). The vertical dashed black lines signify the horizon limit for this analysis using $30$\,m baselines. We highlight that the bootstrapped data points and thermal noise prediction show good agreement, while the limits from the full injection framework (red and blue) are inflated due to the additional inclusion of sample variance that comes from the injection simulations.}
	\label{fig:ps_noise}
\end{figure*}

@@

@@

In Section \ref{sec:BiasOverview} we highlighted some common sources of bias that can show up as power spectrum 
detections and imitate an EoR signal. We discussed the importance of using jackknife and null tests for instilling confidence in an EoR 
detection, as well as identifying other sources of biases. Here we demonstrate methods used by PAPER-64 to mitigate 
foreground and noise bias and we perform null tests in order to characterize the stability and implications of our results.

>>>>

The previously best published 21\,cm upper limit result from \citetalias{ali_et_al2015} placed a $2\sigma$ upper limit 
on $\Delta^{2}(k)$, defined as

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_data.pdf}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data.pdf}
	\caption{A power spectrum of a subset of PAPER-64 data illustrating the use of empirical inverse covariance weighting. The solid red curve is the $2\sigma$ upper limit on the EoR signal estimated from our signal injection framework using empirical inverse covariance weighting. Shown for comparison is the lossy limit prior to signal loss estimation (dashed red). The theoretical $2\sigma$ thermal noise level prediction based on observational parameters is in green, whose calculation is detailed in Section \ref{sec:PSSense}. Additionally, the power spectrum result for the uniform weighted case is shown in three different ways: power spectrum values (black and gray points as positive and negative values, respectively, with $2\sigma$ error bars from bootstrapping), the $2\sigma$ upper limit on the EoR signal using our full signal injection framework (solid blue), and the measured power spectrum values with $2\sigma$ thermal noise errors (gray shaded regions). The vertical dashed black lines signify the horizon limit for this analysis using $30$\,m baselines. In this example, we see that the lossy power spectrum limit is $\sim 4$ orders of magnitude too low when using empirical inverse covariance weighting.}
\label{fig:ps2_data}
\end{figure*}

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_data_add.pdf}
	\includegraphics[width=0.4\textwidth]{plots/ps2_data_add.pdf}
	\caption{A power spectrum of a subset of PAPER-64 data illustrating the use of $\widehat{\textbf{C}}_{\rm eff}$ to minimize signal loss. The solid red curve is the $2\sigma$ upper limit on the EoR signal estimated from our signal injection framework. The theoretical $2\sigma$ thermal noise level prediction based on observational parameters is in green. Additionally, the power spectrum result for the uniform weighted case is shown in three different ways: power spectrum values (black and gray points as positive and negative values, respectively, with $2\sigma$ error bars from bootstrapping), the $2\sigma$ upper limit on the EoR signal using our full signal injection framework (solid blue), and the measured power spectrum values with $2\sigma$ thermal noise errors (gray shaded regions). The vertical dashed black lines signify the horizon limit for this analysis using $30$\,m baselines. This power spectrum result does not use the full dataset's sensitivity as in \citetalias{ali_et_al2015} and Kolopanis et al. (\textit{in prep.}), though we include all analysis changes which have mostly stemmed from revisions regarding signal 
loss, bootstrapping, and the theoretical error computation. We see that the regularization scheme used here produces limits similar to the unweighted limits.}
	\label{fig:ps1_data}
\end{figure*}

@@

In this section we discuss the ways in which we estimate errors for PAPER-64 power spectra. We first walk through an expression for a theoretical error estimation (of thermal noise) based on observational parameters. Although a theoretical model often 
differs from true errors, it is helpful to understand the ideal case and the factors 
that affect its sensitivity. Additionally, we build on the lessons learned about bootstrapping in Section \ref{sec:ErrorOverview} to 
revise our bootstrapping method as applied to PAPER-64 data in order to compute accurate errors from the data itself.

@@

This is lower than the $T_{\rm sys}$ of $500$\,K used in \citetalias{ali_et_al2015} because of several small miscalculation errors that were 

@@

N_{\rm gps}}{2}}$\Big). A revised \citetalias{ali_et_al2015} analysis should therefore use $N_{\rm bls} \sim 32$ instead of $51$, and this change is taken into account in Figure \ref{fig:sense_check}. However, the analysis in this paper and in Kolopanis et al. (\textit{in prep.}) no longer averages baselines into groups ($N_{\rm gps} = 1$). For the subset of data presented in this paper, $N_{\rm bls} = 10$.

@@

An additional factor of $\sqrt{2}$ is gained in sensitivity when folding together positive and negative $k$'s to form $\Delta^{2}(k)$.

@@

\begin{figure}
	\centering
	\includegraphics[width=\columnwidth]{plots/sense_check.pdf}
	\caption{An updated prediction for the thermal noise level of PAPER-64 data (black) is shown in comparison to previously 
published sensitivity limits (gray), both computed for the parameters and methods used in \citetalias{ali_et_al2015}. Major factors that contribute to the discrepancy are $
\Omega_{\rm eff}$, $N_{\rm days}$ and $N_{\rm bls}$, as in Equation \eqref{eq:sense} and described in Section \ref{sec:PSSense}, which when combined decreases our 
sensitivity (higher noise floor) by a factor of $\sim7$ in mK$^{2}$.}
	\label{fig:sense_check}
\end{figure}

@@

\noindent where $\Delta\nu$ is the channel spacing, $\Delta t$ is the integration time, $N_{\rm days}$ is the number of daily counts for a 
particular time and frequency that went into our LST-binned set, and $N_{\rm pols}$ is the number of polarizations ($2$ for Stokes 
I). This temperature sets the variance of the Gaussian random noise.

@@

\begin{figure*}
	\centering
	\includegraphics[width=0.4\textwidth]{plots/ps1_noise_add.pdf}
	\includegraphics[width=0.4\textwidth]{plots/ps2_noise_add.pdf}
	\caption{The power spectrum for a noise simulation that mimics the noise level of a subset of PAPER-64 data, where the solid red curve is the $2\sigma$ upper limit on the EoR signal estimated from our signal injection framework using $\widehat{\textbf{C}}_{\rm eff}$. The theoretical $2\sigma$ thermal noise level prediction based on observational parameters (calculated by Equation \eqref{eq:sense}) is in green. Additionally, the power spectrum result for the uniform weighted case is shown in three different ways: power spectrum values (black and gray points as positive and negative values, respectively, with $2\sigma$ error bars from bootstrapping), the $2\sigma$ upper limit on the EoR signal using our full signal injection framework (solid blue), and the measured power spectrum values with $2\sigma$ thermal noise errors (gray shaded regions). The vertical dashed black lines signify the horizon limit for this analysis using $30$\,m baselines. We highlight that the bootstrapped data points and thermal noise prediction show good agreement, while the limits from the full injection framework (red and blue) are inflated due to the additional inclusion of sample variance that comes from the injection simulations.}
	\label{fig:ps_noise}
\end{figure*}

@@

In summary, we have learned several lessons regarding bootstrapping and have revised our analysis procedure in order to determine error bars that correctly reflect the variance in our power spectrum estimates. Additionally, we have carried out substantial testing with these changes in place to verify that the computed errors for our noise simulations are accurate and agree with thermal noise predictions.

@@

In Section \ref{sec:BiasOverview} we highlighted some common sources of bias that can show up as power spectrum 
detections and imitate an EoR signal. We discussed the importance of using jackknife and null tests for instilling confidence in an EoR 
detection, as well as for identifying other sources of biases. Here we demonstrate methods used by PAPER-64 to mitigate 
foreground and noise bias and we perform null tests in order to characterize the stability and implications of our results.
